Application of Generalized Additive Models for long-term and seasonal trend analysis of water quality: South San Francisco Bay case study

### 2020-06-15 manu_draft.Rmd
new model summary table with log and boxcox

```{r modprftab, tab.cap = 'Model performance statistics for each station and GAM structure (Table \\@ref(tab:modsumtab)).  Four GAMs were fit to each station and performance is summarized as $\\Delta$ AIC (from `gam0`), GCV, and R-squared values.  Models with the lowest AIC, GCV and highest R-squared values for each station are emphasized. Probability values ("P-value") show the results of Analysis of Variance comparisons between all models at a station.  The p-value for a row refers to the comparison of the model in the current row to the one preceeding (`gam0` is the base model and does not have a p-value).  A significant p-value indicates the model in the current row provides a statistically significant difference in explaining chlorophyll variance from the model in the preceding row.  Values indicate ns for not significantly different at $\\alpha = 0.05$, * $p< 0.05$, and ** $p < 0.005$.'}
totab <- modprf %>% 
  group_by(station) %>% 
  mutate(
    AIC = AIC - max(AIC),
    minaic = AIC == min(AIC), 
    mingcv = GCV == min(GCV),
    maxr2 = R2 == max(R2)
  ) %>% 
  ungroup

@@

>>>>

```{r modprftab, tab.cap = 'Model performance statistics for each station and GAM structure (Table \\@ref(tab:modsumtab)). Results are for models of chlorophyll-a using BoxCox and logarathmic transformations. Four GAMs were fit to each station and performance is summarized as $\\Delta$ AIC (from `gam0`),  $\\Delta$ GCV, and R-squared values.  Probability values ("P-val") show the results of Analysis of Variance comparisons between all BoxCox or logarithmic models at a station.  The p-value for a row refers to the comparison of the model in the current row to the one preceeding (`gam0` is the base model and does not have a p-value).  A significant p-value indicates the model in the current row provides a statistically significant difference in explaining chlorophyll variance from the model in the preceding row.  ns: not significant at $\\alpha = 0.05$, * $p< 0.05$, and ** $p < 0.005$.'}

totab <- modprf %>% 
  gather('stat', 'val', AIC, GCV, R2, `P-value`) %>% 
  mutate(trans = case_when(
    trans != 'log10' ~ 'BoxCox', 
    T ~ trans)
  ) %>% 
  unite('stat', trans, stat) %>% 
  mutate(
    stat = factor(stat, levels = c('BoxCox_AIC', 'log10_AIC', 'BoxCox_GCV', 'log10_GCV', 'BoxCox_R2', 'log10_R2', 'BoxCox_P-value', 'log10_P-value'))
  ) %>% 
  spread(stat, val) %>% 
  mutate_at(c('BoxCox_AIC', 'log10_AIC', 'BoxCox_GCV', 'log10_GCV', 'BoxCox_R2', 'log10_R2'), as.numeric) %>% 
  group_by(station) %>%
  mutate(
    BoxCox_AIC = BoxCox_AIC - max(BoxCox_AIC),
    log10_AIC = log10_AIC - max(log10_AIC),
    BoxCox_GCV = BoxCox_GCV - max(BoxCox_GCV),
    log10_GCV = log10_GCV - max(log10_GCV)#,
    # minaic = AIC == min(AIC), 
    # mingcv = GCV == min(GCV),
    # maxr2 = R2 == max(R2)
  ) %>% 
  ungroup

# boldaic <- as.numeric(rownames(totab)[totab$minaic])
# boldgcv <- as.numeric(rownames(totab)[totab$mingcv])
# boldr2 <- as.numeric(rownames(totab)[totab$maxr2])

totab <- totab %>% 
  # select(-minaic, -mingcv, -maxr2) %>% 
  mutate(
    station = ifelse(duplicated(station), '', station)
  )

tab <- flextable(totab) %>% 
  set_header_labels(
    station = '', 
    model = '', 
    BoxCox_AIC = 'Delta AIC',
    log10_AIC = 'Delta AIC',
    BoxCox_GCV = 'Delta GCV',
    log10_GCV = 'Delta GCV',
    BoxCox_R2 = 'R-squared',
    log10_R2 = 'R-squared',
    `BoxCox_P-value` = 'P-val',
    `log10_P-value` = 'P-val'
  ) %>% 
  merge_at(i = 1, j = 3:4, part = "header") %>% 
  merge_at(i = 1, j = 5:6, part = "header") %>% 
  merge_at(i = 1, j = 7:8, part = "header") %>% 
  merge_at(i = 1, j = 9:10, part = "header") %>% 
  add_header_row(values = c('Station', 'GAM', 'BoxCox', 'Log', 'BoxCox', 'Log', 'BoxCox', 'Log', 'BoxCox', 'Log'), top = F) %>% 
  align(j = c(3:10), align = 'center', part = 'all') %>% 
  colformat_num(j = c(3, 4), digits = 1) %>%
  colformat_num(j = c(5, 6), digits = 3) %>% 
  colformat_num(j = c(7, 8), digits = 2) %>% 
  # bold(i = boldaic, j = 'AIC') %>% 
  # bold(i = boldgcv, j = 'GCV') %>% 
  # bold(i = boldr2, j = 'R2') %>% 
  border_remove() %>% 
  hline_top(border = fp_border(color = 'black')) %>% 
  font(fontname = 'Times', part = 'all') %>%
  font(j = 2, fontname = 'Consolas') %>% 
  fontsize(size = 8, part = 'all') %>% 
  autofit()

# Supplement

```{r modprfboxtab, tab.cap = 'Model performance statistics for each station and GAM structure (Table \\@ref(tab:modsumtab)). Results are for models of chlorophyll-a using BoxCox transformations. Four GAMs were fit to each station and performance is summarized as $\\Delta$ AIC (from `gam0`), GCV, and R-squared values.  Models with the lowest AIC, GCV and highest R-squared values for each station are emphasized. Probability values ("P-value") show the results of Analysis of Variance comparisons between all models at a station.  The p-value for a row refers to the comparison of the model in the current row to the one preceeding (`gam0` is the base model and does not have a p-value).  A significant p-value indicates the model in the current row provides a statistically significant difference in explaining chlorophyll variance from the model in the preceding row.  Values indicate ns for not significantly different at $\\alpha = 0.05$, * $p< 0.05$, and ** $p < 0.005$.'}
totab <- modprf %>% 
  filter(trans != 'log10') %>% 
  select(-trans) %>% 
  group_by(station) %>% 
  mutate(
    AIC = AIC - max(AIC),
    minaic = AIC == min(AIC), 
    mingcv = GCV == min(GCV),
    maxr2 = R2 == max(R2)
  ) %>% 
  ungroup

@@

```{r modprflogtab, tab.cap = 'Model performance statistics for each station and GAM structure (Table \\@ref(tab:modsumtab)). Results are for models of chlorophyll-a using logarithmic (base-10) ransformations. Four GAMs were fit to each station and performance is summarized as $\\Delta$ AIC (from `gam0`), GCV, and R-squared values.  Models with the lowest AIC, GCV and highest R-squared values for each station are emphasized. Probability values ("P-value") show the results of Analysis of Variance comparisons between all models at a station.  The p-value for a row refers to the comparison of the model in the current row to the one preceeding (`gam0` is the base model and does not have a p-value).  A significant p-value indicates the model in the current row provides a statistically significant difference in explaining chlorophyll variance from the model in the preceding row.  Values indicate ns for not significantly different at $\\alpha = 0.05$, * $p< 0.05$, and ** $p < 0.005$.'}
totab <- modprf %>% 
  filter(trans == 'log10') %>% 
  select(-trans) %>% 
  group_by(station) %>% 
  mutate(
    AIC = AIC - max(AIC),
    minaic = AIC == min(AIC), 
    mingcv = GCV == min(GCV),
    maxr2 = R2 == max(R2)
  ) %>% 
  ungroup
